---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mvpd: An R Package for Multivariate Product Distributions

  * `[dpr]mvss`: mutivariate subgaussian stable distributions
  
<!-- badges: start -->
<!-- badges: end -->

The goal of `mvpd` is to use product distribution theory to allow the 
numerical calculations of specific scale mixtures of the multivariate normal
distribution.  The multivariate subgaussian stable distribution is the product
of the square root of a positive stable distribution and the multivariate normal 
distribution (see Nolan (2013)).

## Installation

You can install the development version of mvpd from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("swihart/mvpd")
```

## Example

Generate 1000 draws from a random bivariate subgaussian stable distribution
with alpha=1.71 and plot.

```{r example, message=FALSE}
library(mvpd)
## basic example code
biv <- rmvss(n=1e3, alpha=1.71, Q=matrix(c(10,7.5,7.5,10),2))
head(biv)
plot(biv); abline(h=0,v=0)
```
