---
title: "Introduction and Primer"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

## https://stackoverflow.com/questions/66011929/package-rgl-in-r-not-loading-in-mac-os/66127391#66127391
options(rgl.useNULL = TRUE)
library(rgl)

```

```{r setup}
library(uniformly)
set.seed(6789)
## creator view
nsim <- 1e3
d <- 2
chisq_draw <- rchisq(nsim, df=d)
#chisq_draw <- rexp(nsim, rate=1)
# chisq_draw <- 4
# chisq_draw <- sample(c(4,1,0), nsim, TRUE)
# chisq_draw <- runif(nsim, 0, 4)
# chisq_draw <- -1
# chisq_draw <- rbinom(nsim,1,prob=0.5 )
# chisq_draw <- rpois(nsim,1)
# chisq_draw <- rbeta(nsim,.20,.20)
# chisq_draw <- rbeta(nsim,2.0,2.0)
# chisq_draw <- rbeta(nsim,20,20)
# chisq_draw <- rgamma(nsim, 1,.1)
#chi_sq_draw <- rcauchy(nsim,0,10)^2 ## see connection to betaprime
```

Take the draw from above.


Apply it to the uniform on the sphere (for d=2 this is a circle).

```{r}
unif_draw <- uniformly::runif_on_sphere(nsim,d)
```

Get a nice rotation going with chol.

```{r}
spread <- 1
rho <- -0.80
SIGMA <- spread*diag(d) + rho - diag(d)*rho
SIGMA
chol(SIGMA)
```

Out of all the stuff generated, find the max so we can standardize the axes.

```{r}
maxrange <- max(c(sqrt(abs(chisq_draw)) * unif_draw %*% chol(SIGMA),
                                          unif_draw %*% chol(SIGMA)
                  )
)


```

This is the uniform circle.

```{r}
par(pty='s', mfrow=c(2,2), bty="n", oma=rep(0,4), mai=c(0.3,0,0.3,0))
plot( unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main=expression(paste("U")))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot( unif_draw %*% chol(SIGMA), 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main=expression(paste("U", Sigma^{~~1/2})))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot(sign(chisq_draw)*sqrt(abs(chisq_draw)) * unif_draw %*% chol(SIGMA), 
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main=expression(paste(A^{1/2},"U", Sigma^{~~1/2})))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


par(pty='s')
plot(mvtnorm::rmvnorm(nsim, sigma=SIGMA),
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main="rnorm(nsim, 0, SIGMA)")
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")



```

$U$ is the uniform circle transformed into an ellipse $U \Sigma^{1/2}$, where $\Sigma^{1/2}$ = (R-code: `chol(SIGMA)`).

Then each point of the ellipse is multiplied by the square root of the random draw from the mixing distribution $A$.

And a MVN(0, Sigma) draw for comparison, as procured by `rnorm(nsim,0,SIGMA)`.
