---
title: "Introduction and Primer II"
subtitle: Start with a unit circle.  End with a Bivariate Normal.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

## https://stackoverflow.com/questions/66011929/package-rgl-in-r-not-loading-in-mac-os/66127391#66127391
options(rgl.useNULL = TRUE)
library(rgl)

```

```{r setup, include=FALSE}
library(uniformly)
set.seed(6789)
## creator view
nsim <- 1e3
d <- 2
#chisq_draw <- rchisq(nsim, df=d)
#chisq_draw <- rexp(nsim, rate=1)
# chisq_draw <- 4
# chisq_draw <- sample(c(4,1,0), nsim, TRUE)
# chisq_draw <- runif(nsim, 0, 4)
# chisq_draw <- -1
# chisq_draw <- rbinom(nsim,1,prob=0.5 )
# chisq_draw <- rpois(nsim,1)
# chisq_draw <- rbeta(nsim,.20,.20)
# chisq_draw <- rbeta(nsim,2.0,2.0)
# chisq_draw <- rbeta(nsim,20,20)
# chisq_draw <- rgamma(nsim, 1,.1)
#chi_sq_draw <- rcauchy(nsim,0,10)^2 ## see connection to betaprime

# chisq_draw <- runif(nsim, 0.6, 10)
# chisq_draw <- sample(c(0.6,10), nsim, TRUE)

alpha <- 1.50
# chisq_draw <- libstable4u::stable_rnd(nsim,
#                                        pars=c(
#                                          alpha = alpha/2,
#                                          beta = 1,
#                                          sigma = 2*cos(pi*alpha/4)^(2/alpha),
#                                          mu = 0),
#                                        parametrization = 1L)
# 
# 
# chisq_draw <- libstable4u::stable_rnd(nsim,
#                                        pars=c(
#                                          alpha = alpha/2,
#                                          beta = 1,
#                                          sigma = 2*cos(pi*alpha/4)^(2/alpha),
#                                          mu = 0),
#                                        parametrization = 1L) *
#               rchisq(nsim, df=d)
# 
# 
# 

chisq_draw <- stabledist::rstable(nsim,
                                         alpha = alpha/2,
                                         beta = 1,
                                         gamma = 2*cos(pi*alpha/4)^(2/alpha),
                                         delta = 0,
                                       pm = 1L)
# chisq_draw <- stabledist::rstable(nsim,
# 
#                                          alpha = alpha/2,
#                                          beta = 1,
#                                          gamma = 2*cos(pi*alpha/4)^(2/alpha),
#                                          delta = 0,
#                                        pm = 1L)*
#   rchisq(nsim, df=d)

# 


```


```{r, include=FALSE}
## Apply it to the uniform on the sphere (for d=2 this is a circle).
##set.seed(6789)
unif_draw <- uniformly::runif_on_sphere(nsim,d)
## Get a nice rotation going with chol.
spread <- 1
rho <- -0.80
SIGMA <- spread*diag(d) + rho - diag(d)*rho
SIGMA
chol(SIGMA)
```


```{r, include=FALSE}
## Out of all the stuff generated, find the max so we can standardize the axes.

simmed_mvnorm <- mvtnorm::rmvnorm(nsim, sigma=SIGMA)
simmed_mvnorm <- mvpd::rmvss(nsim, Q=SIGMA, alpha=alpha)
maxrange <- max(abs(c(sqrt(abs(chisq_draw)) * unif_draw %*% chol(SIGMA),
                                          unif_draw %*% chol(SIGMA),
                  c(simmed_mvnorm)
                  ))
)

maxrange <- c(5)
```



```{r,echo=FALSE, fig.width=8, fig.height=8}
par(pty='s', mfrow=c(2,2), bty="n", oma=rep(0,4), mai=c(0.3,0,0.3,0))
plot( unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d])))
      main=expression(paste(U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot( sign(chisq_draw)*sqrt(abs(chisq_draw)) *unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d], Sigma^{~~1/2})))
      main=expression(paste(B, U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot( sign(chisq_draw)*sqrt(abs(chisq_draw)) *unif_draw %*% chol(SIGMA), 
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(A^{1/2}, U[d], Sigma^{~~1/2})))
     ##main=expression(paste(B[1], U[d], R[d])))
     main=expression(paste(B, U, R)))
      # main="BUR")
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


par(pty='s')
plot(simmed_mvnorm,
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main="rnorm(nsim, 0, Q)")
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


# 
# `r expression(paste(U))` is the uniform circle transformed into an ellipse `r expression(paste(U,R))`, where `r expression(R)`=$\Sigma^{1/2}$ = (R-code: `chol(SIGMA)`).
# 
# Then each point of the ellipse is multiplied by `r expression(paste(B))`= `r expression(paste(A^{1/2}))`, the square root of the random draw from distribution `r expression(paste(A))`.
# 
# And a MVN(0, Sigma) draw for comparison, as procured by `rnorm(nsim,0,SIGMA)`.

```


  * `r expression(U)` is a nx2 matrix containing n random draws from the uniform circle.
  * B = sqrt(A), where A is a nx1 matrix containing n *i.i.d* draws from its univariate distribution.
  * `r expression(BU)` can be thought of as the distribution of radii
  * `r expression(BUR)` is the result of the matrix multiplication between nx2 `r expression(BU)` and 2x2 `r expression(R)`, the Cholesky decomposition of a 2x2 matrix `r expression(Q)`.
    * R makes a circular (`r expression(BU)`) into an elliptical (`r expression(BUR)`)
    * R can be thought of as a "square root" of a matrix. `r expression(R)`'`r expression(R)` = `r expression(Q)`.
  * When A $\sim \chi^2$(df=2) distribution, BUR is the bivariate normal distribution G(0,Q). Equivalently,
  * When B $\sim \chi$(df=2) distribution, BUR is the bivariate normal distribution G(0,Q)



<!-- Graveyard / density of code? -->
  <!-- you used this to  trouble shoot the RNG problem -->
  <!-- ultimately found in libstable4u -->
  
```{r, echo=FALSE, eval=FALSE, include=FALSE}
runif_on_sphere2 <- function(n, d, r=1){
  sims <- matrix(rnorm(n*d), nrow=n, ncol=d)
  r * sims / sqrt(apply(sims, 1L, crossprod))
}

set.seed(67890)
d <- 2
nsim <- 1e3
alpha <- 1.2
draw <- libstable4u::stable_rnd(nsim,
                                pars=c(
                                  alpha = alpha/2,
                                  beta = 1,
                                  sigma = 2*cos(pi*alpha/4)^(2/alpha),
                                  mu = 0),
                                parametrization = 1L)



xdom <- c(seq(0.0,1,0.01), seq(1.1,6,.1))
yden <-  2*xdom*libstable4u::stable_pdf(xdom^2,
                                        pars=c(
                                          alpha = alpha/2,
                                          beta = 1,
                                          sigma = 2*cos(pi*alpha/4)^(2/alpha),
                                          mu = 0),
                                        parametrization = 1L)

hist(draw[draw<6], freq=FALSE,breaks=100)

plot(xdom,yden)
hist(draw[draw<36]^(1/2), freq=FALSE, breaks=100)
min(draw^(1/2))
max(draw^(1/2))
#plot(ecdf(sqrt(draw)), xlim=c(0,2), ylim=c(0,0.4))


#set.seed(1)
#unif_draw <- uniformly::runif_on_sphere(nsim,d)
unif_draw <-runif_on_sphere2(nsim,d)

tol <- 1e-2

head(unif_draw[unif_draw[,1] < tol & unif_draw[,1] > -tol, ])
head(sqrt(draw)[unif_draw[,1] < tol & unif_draw[,1] > -tol])
head(sqrt(draw)[unif_draw[,1] < tol & unif_draw[,1] > -tol]*unif_draw[unif_draw[,1] < tol & unif_draw[,1] > -tol, ])

head(unif_draw[unif_draw[,2] < tol & unif_draw[,2] > -tol, ])
head(sqrt(draw)[unif_draw[,2] < tol & unif_draw[,2] > -tol])
head(sqrt(draw)[unif_draw[,2] < tol & unif_draw[,2] > -tol]*unif_draw[unif_draw[,2] < tol & unif_draw[,2] > -tol, ])



par(pty='s', mfrow=c(1,2), bty="n", oma=rep(0,4), mai=c(0.3,0,0.3,0))
plot( unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d])))
      main=expression(paste(U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot( sqrt(draw) *unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d], Sigma^{~~1/2})))
      main=expression(paste(B, U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


par(pty='s')
plot( sign(draw)*sqrt(abs(draw)) *unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d], Sigma^{~~1/2})))
      main=expression(paste(B, U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


```