---
title: "Introduction and Primer"
subtitle: Start with a unit circle.  End with a Bivariate Normal.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

## https://stackoverflow.com/questions/66011929/package-rgl-in-r-not-loading-in-mac-os/66127391#66127391
options(rgl.useNULL = TRUE)
library(rgl)

```

```{r setup, include=FALSE}
library(uniformly)
set.seed(6789)
## creator view
nsim <- 1e3
d <- 2
chisq_draw <- rchisq(nsim, df=d)
#chisq_draw <- rexp(nsim, rate=1)
# chisq_draw <- 4
# chisq_draw <- sample(c(4,1,0), nsim, TRUE)
# chisq_draw <- runif(nsim, 0, 4)
# chisq_draw <- -1
# chisq_draw <- rbinom(nsim,1,prob=0.5 )
# chisq_draw <- rpois(nsim,1)
# chisq_draw <- rbeta(nsim,.20,.20)
# chisq_draw <- rbeta(nsim,2.0,2.0)
# chisq_draw <- rbeta(nsim,20,20)
# chisq_draw <- rgamma(nsim, 1,.1)
#chi_sq_draw <- rcauchy(nsim,0,10)^2 ## see connection to betaprime
```


```{r, include=FALSE}
## Apply it to the uniform on the sphere (for d=2 this is a circle).
##set.seed(1)
unif_draw <- uniformly::runif_on_sphere(nsim,d)
## Get a nice rotation going with chol.
spread <- 1
rho <- -0.80
SIGMA <- spread*diag(d) + rho - diag(d)*rho
SIGMA
chol(SIGMA)
```


```{r, include=FALSE}
## Out of all the stuff generated, find the max so we can standardize the axes.


simmed_mvnorm <- mvtnorm::rmvnorm(nsim, sigma=SIGMA)
maxrange <- max(c(sqrt(abs(chisq_draw)) * unif_draw %*% chol(SIGMA),
                                          unif_draw %*% chol(SIGMA),
                  c(simmed_mvnorm)
                  )
)



```



```{r,echo=FALSE, fig.width=8, fig.height=8}
par(pty='s', mfrow=c(2,2), bty="n", oma=rep(0,4), mai=c(0.3,0,0.3,0))
plot( unif_draw, 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d])))
      main=expression(paste(U)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot( unif_draw %*% chol(SIGMA), 
      xlim=maxrange*c(-1,1),
      ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(U[d], Sigma^{~~1/2})))
      main=expression(paste(U, A)))
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")

par(pty='s')
plot(sign(chisq_draw)*sqrt(abs(chisq_draw)) * unif_draw %*% chol(SIGMA), 
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      ##main=expression(paste(A^{1/2}, U[d], Sigma^{~~1/2})))
     ##main=expression(paste(B[1], U[d], R[d])))
     main=expression(paste(R, U, A)))
      # main="BUR")
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


par(pty='s')
plot(simmed_mvnorm,
     xlim=maxrange*c(-1,1),
     ylim=maxrange*c(-1,1),
      xlab="",
      ylab="",
      main="rnorm(nsim, 0, Q)")
abline(h=c(-1,1), v=c(-1,1), lty=2, col="lightgrey")


# 
# `r expression(paste(U))` is the uniform circle transformed into an ellipse `r expression(paste(U,R))`, where `r expression(R)`=$\Sigma^{1/2}$ = (R-code: `chol(SIGMA)`).
# 
# Then each point of the ellipse is multiplied by `r expression(paste(B))`= `r expression(paste(A^{1/2}))`, the square root of the random draw from distribution `r expression(paste(A))`.
# 
# And a MVN(0, Sigma) draw for comparison, as procured by `rnorm(nsim,0,SIGMA)`.
```
```{r,eval=FALSE,include=FALSE}
mean(simmed_mvnorm[,1] < 1 & 
     simmed_mvnorm[,1] > -1 & 
     simmed_mvnorm[,2] < 1 & 
     simmed_mvnorm[,2] > -1
)

BUR <- sqrt(chisq_draw) * unif_draw %*% chol(SIGMA)

mean(BUR[,1] < 1 & 
     BUR[,1] > -1 & 
     BUR[,2] < 1 & 
     BUR[,2] > -1
)

verify_pmvnorm <- mvtnorm::pmvnorm(lower=c(-1,-1), upper=c(1,1), sigma=SIGMA)
verify_pmvnorm

mean(simmed_mvnorm[,1] < Inf & 
     simmed_mvnorm[,1] > 1 & 
     simmed_mvnorm[,2] < 1 & 
     simmed_mvnorm[,2] > -1
)

BUR <- sqrt(chisq_draw) * unif_draw %*% chol(SIGMA)

mean(BUR[,1] < Inf & 
     BUR[,1] > 1 & 
     BUR[,2] < 1 & 
     BUR[,2] > -1
)

verify_pmvnorm <- mvtnorm::pmvnorm(lower=c(1,-1), upper=c(Inf,1), sigma=SIGMA)
verify_pmvnorm

```


  * `r expression(U)` is a `n x 2` matrix containing n random draws from the uniform circle.
  * `r expression(UA)` is the result of the matrix multiplication between `n x 2` `r expression(U)` and `2 x 2` `r expression(A)`, the Cholesky decomposition of a `2x2` matrix `r expression(Q)`.
    * A makes a circle (`r expression(U)`) into an ellipse (`r expression(UA)`)
    * A can be thought of as a "square root" of a matrix. `r expression(A)`'`r expression(A)` = `r expression(Q)`.
  * When R $\sim \chi$(df=2) distribution, RUA is the bivariate normal distribution G(0,Q), Equivalently
    * `mvpd` uses G for Gaussian
