---
title: "(XY)^Z Part I"
author: Bruce J. Swihart
date: 2025-05-16
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(mvpd)
```

___ 

**IN PROGRESS**

---

I'm about 8 months late to the party, but a challenge problem from 3blue1brown
caught my attention, as well as a call for intuitive approaches.

> Here's the challenge mode for all you math whizzes. Sample three numbers x, y, z uniformly at random in [0, 1], and compute (xy)^z. What distribution describes this result?
>
> Answer: It's uniform!
>
> I know how to prove it, but haven't yet found the "aha" style explanation where it feels expected or visualizable. If any of you have one, please send it my way and I'll strongly consider making a video on it.
>
> -- Grant Sanderson of 3blue1brown, 2024-09-10


The result and good explanations of the result have been around for a while, as shown in this [Math SE post from 2012](https://math.stackexchange.com/q/261783/180716).  



## Bonus: $\alpha_x \neq \alpha_y$

  * [Wikipedia Sum of Two Independent Exponential Random Variables](https://en.wikipedia.org/wiki/Exponential_distribution#Sum_of_two_independent_exponential_random_variables)