---
title: "(XY)^Z Part II"
author: Bruce J. Swihart
date: 2025-05-16
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
nsim.switch <- 1e4
overlaid.dens.lwd <- 3
## https://stackoverflow.com/a/57157075/2727349
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", 
                             "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", 
                             "#661100", "#6699CC", "#888888")
scales::show_col(safe_colorblind_palette)


```

```{r setup, include=FALSE}
library(mvpd)
```

___ 

**IN PROGRESS**

---


So far we have considered cases where the alpha for X and Y are the same and how multiplying (log(1/X)+log(1/Y)) by Z allows one to recover the original distribution of log(1/X) (equivalently, log(1/Y)).  However, if $\alpha_x \neq \alpha_y$, multiplying (log(1/X)+log(1/Y)) by a uniform would not recover the distribution for either log(1/X) or log(1/Y) -- it would be a distribution "in-between" those two distributions -- but not an Exponential with an alpha between $\alpha_x$ and $\alpha_y$.  We consider this "unequal alpha" as a bonus case below.


## Bonus: Unequal alpha cases

Firstly, one may want to consult Wikipedia.

  * [Wikipedia Sum of Two Independent Exponential Random Variables](https://en.wikipedia.org/wiki/Exponential_distribution#Sum_of_two_independent_exponential_random_variables)
  

So, in this setup, there is no *one* distribution to recover since  $\alpha_x \neq \alpha_y$, so trying different distributions for $Z$ leads to different blends of the X,Y distributions for the quantity Z(log(1/X)+log(1/Y)).  

### Z distributed as U(0,1)

  * $Z \sim$ U(0,1)
  <!-- * mathematica says I need incomplete gamma. luckily [vincent goulet made one for R](https://search.r-project.org/CRAN/refmans/expint/html/gammainc.html). -->
  <!-- * page 80 green notebook  -->
  <!-- * describe what each color density is -->
  * magenta density
    * $f_w(w) = - \frac{mn \left( \Gamma(0,mw) - \Gamma(0,nw)\right)}{ m-n }$ where $W = Z \left( \log \frac{1}{X} + \log \frac{1}{Y} \right)$, m=$\alpha_x$, n=$\alpha_y$
    * $\Gamma(.,.)$ is [incomplete Gamma function](https://en.wikipedia.org/wiki/Incomplete_gamma_function#Special_values).

```{r, results='hide', echo=FALSE, warning=FALSE}
## sum of two indep exponentials aa_x != aa_y
## Z is uniform
nsims <- nsim.switch
aa_x <- 4
aa_y <- 2

bb_x <- 1
bb_y <- 1

x <- rbeta(nsims, aa_x, bb_x)
y <- rbeta(nsims, aa_y, bb_y)
z <- runif(nsims, 0, 1)


s.xlim <- range(c(log(1/x),
                  log(1/y),
                  z,
                  log(1/x * 1/y),
                  z*log(1/x*1/y)
)
) #c(-10,0)#c(-4,2)

s.xlim <- c(0,3)

s.ylim <- c(0, max(dbeta(seq(s.xlim[1],s.xlim[2],0.01), aa_x, bb_x)))
s.ylim
s.xlim

#dlogbeta<-function(x,a,b) exp(a*x) * (1-exp(x))^(b-1) / beta(a,b)

par(mfcol=c(3,3), mai=c(0.3,0.3,0.3,0.3))
hist(log(1/x), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_x),
      lwd=overlaid.dens.lwd,
      col="blue"
)

hist(log(1/y), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/Y)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_y),
      lwd=overlaid.dens.lwd,
      col="cyan"
)

hist(    z , freq=FALSE, breaks=50, ylim=s.ylim, xlim=c(0,1), 
     main="Z", xlab="",ylab="")




plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
hist(log(1/x)+log(1/y), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)+log(1/Y)", xlab="",ylab="")
f_1x_1y <- function(w, m, n){
  
  m*n/(n-m) * (exp(-m*w) - exp(-n*w))
  
}
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      f_1x_1y(seq(s.xlim[1], s.xlim[2],0.01 ),
              aa_x,
              aa_y),
      lwd=overlaid.dens.lwd,
      col="darkorange"
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")

plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
hist(z*(log(1/x)+log(1/y)), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="Z(log(1/X)+log(1/Y))", xlab="",ylab="")


fw_dens <- function(w, m, n){
  
  num = m*n*(expint::gammainc(0,m*w) - expint::gammainc(0, n*w))
  den = m-n
  
  -num/den
  
}

lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      fw_dens(seq(s.xlim[1], s.xlim[2],0.01 ),
              aa_x,
              aa_y),
      lwd=overlaid.dens.lwd,
      col="magenta"
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")


```

  * While the computed quantity is not exponential, one can see it is a concentration of the convolution and that it is kind of "in-between" the two component distributions.
  
### Z distributed as Inverse-Gamma(alpha, beta)

  * $Z \sim$ Inverse-Gamma($\alpha_z$, $\beta_z$)
  <!-- * page 83 green notebook -->
  <!-- * as $\alpha_z = \beta_z >>>> 1$ Z converges on 1 and the distributions look more similar  -->
  <!-- * describe what each color of density is -->
  * gold density
  
    * $f_w(w) = \frac{mn}{ n-m }\frac{\beta_z^{\alpha_z}\Gamma(\alpha_z + 1)}{ \Gamma(\alpha_z) } \left( \frac{1}{ \left(mw+\beta_z\right)^{\alpha_z +1}}-\frac{1}{ \left(nw+\beta_z\right)^{\alpha_z +1}} \right)$ where $W = Z \left( \log \frac{1}{X} + \log \frac{1}{Y} \right)$, m=$\alpha_x$, n=$\alpha_y$.


    
```{r, results='hide', echo=FALSE}
## sum of two indep exponentials aa_x != aa_y
## and z ~ IG(a_z, b_z)

nsims <- nsim.switch
aa_x <- 4
aa_y <- 2

bb_x <- 1
bb_y <- 1

aa_z <- 3
bb_z <- .9

x <- rbeta (nsims, aa_x, bb_x)
y <- rbeta (nsims, aa_y, bb_y)
z <- invgamma::rinvgamma(nsims, shape=aa_z, rate=bb_z)


s.xlim <- c(0,3)
#   range(c(log(1/x),
#                   log(1/y),
#                   z,
#                   log(1/x * 1/y),
#                   z*log(1/x*1/y)
# )) 
#c(-10,0)
#c(-4,2)

s.ylim <- c(0, max(dbeta(seq(s.xlim[1],s.xlim[2],0.01), aa_x, bb_x)))
s.ylim
s.xlim

#dlogbeta<-function(x,a,b) exp(a*x) * (1-exp(x))^(b-1) / beta(a,b)

par(mfcol=c(3,3), mai=c(0.3,0.3,0.3,0.3))
hist(log(1/x), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_x),
      lwd=overlaid.dens.lwd,
      col="blue"
)


hist(log(1/y), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/Y)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_y),
      lwd=overlaid.dens.lwd,
      col="cyan"
)

hist(    z , freq=FALSE, breaks=500, ylim=s.ylim, xlim=c(0,4), 
     main="Z", xlab="",ylab="")


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
hist(log(1/x)+log(1/y), freq=FALSE, breaks=500, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)+log(1/Y)", xlab="",ylab="")

f_1x_1y <- function(w, m, n){
  
  m*n/(n-m) * (exp(-m*w) - exp(-n*w))
  
}

lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      f_1x_1y(seq(s.xlim[1], s.xlim[2],0.01 ),
              aa_x,
              aa_y),
      lwd=overlaid.dens.lwd,
      col="darkorange"
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")

plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")

hist(z*(log(1/x)+log(1/y)), freq=FALSE, breaks=1000, ylim=s.ylim, xlim=s.xlim, 
     main="Z(log(1/X)+log(1/Y))", xlab="",ylab="")


fw_dens_ig <- function(w, m, n, o, p){
  
  fac1 <- (m*n) / (n-m)
  fac2 <- p^o * gamma(o+1) / gamma(o)
  fac3 <- (1/(m*w+p)^(o+1) - 1/(n*w+p)^(o+1))
  
  fac1*fac2*fac3
  
}

lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      fw_dens_ig(seq(s.xlim[1], s.xlim[2],0.01 ),
              aa_x,
              aa_y,
              aa_z,
              bb_z),
      lwd=overlaid.dens.lwd,
      col="#F0E442" 
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
```

  * While the computed quantity is not exponential, one can see it is a concentration of the convolution and that it is kind of "in-between" the two component distributions.
  
### Z distributed as Pareto  

  * $Z \sim$ Pareto($x_m=1$,$\alpha_m=1$) (where 1/Z ~ U(0,1))
  <!-- * page 81 green notebook -->
  <!-- * describe what each color of density is -->

  * magenta density:  
    * $f_w(w) = \frac{n^2 (1-\exp^{-mw}(1+mw)) - m^2 (1-\exp^{-nw}(1+nw))}{ mn(n-m)w^2 }$ where $W = Z \left( \log \frac{1}{X} + \log \frac{1}{Y} \right)$, m=$\alpha_x$, n=$\alpha_y$.
  
```{r, results='hide', echo=FALSE}
## sum of two indep exponentials aa_x != aa_y
## and z ~ IG(a_z, b_z)

nsims <- nsim.switch
aa_x <- 4
aa_y <- 2

bb_x <- 1
bb_y <- 1

aa_z <- 3
bb_z <- .9

x <- rbeta (nsims, aa_x, bb_x)
y <- rbeta (nsims, aa_y, bb_y)
z <- LNPar::rpareto(nsims,1,alpha=1)


s.xlim <- c(0,3)
#   range(c(log(1/x),
#                   log(1/y),
#                   z,
#                   log(1/x * 1/y),
#                   z*log(1/x*1/y)
# )) 
#c(-10,0)
#c(-4,2)

s.ylim <- c(0, max(dbeta(seq(s.xlim[1],s.xlim[2],0.01), aa_x, bb_x)))
s.ylim
s.xlim

#dlogbeta<-function(x,a,b) exp(a*x) * (1-exp(x))^(b-1) / beta(a,b)

par(mfcol=c(3,3), mai=c(0.3,0.3,0.3,0.3))
hist(log(1/x), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_x),
      lwd=overlaid.dens.lwd,
      col="blue"
)


hist(log(1/y), freq=FALSE, breaks=50, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/Y)", xlab="",ylab="")
lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      dexp(seq(s.xlim[1], s.xlim[2],0.01 ),
           aa_y),
      lwd=overlaid.dens.lwd,
      col="cyan"
)

hist(    z , freq=FALSE, breaks=1e5, ylim=c(0,1), xlim=c(0,10), 
     main="Z", xlab="",ylab="")


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
hist(log(1/x)+log(1/y), freq=FALSE, breaks=500, ylim=s.ylim, xlim=s.xlim, 
     main="log(1/X)+log(1/Y)", xlab="",ylab="")

f_1x_1y <- function(w, m, n){
  
  m*n/(n-m) * (exp(-m*w) - exp(-n*w))
  
}

lines(seq(s.xlim[1], s.xlim[2],0.01 ),
      f_1x_1y(seq(s.xlim[1], s.xlim[2],0.01 ),
              aa_x,
              aa_y),
      lwd=overlaid.dens.lwd,
      col="darkorange"
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")

plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")

hist(z*(log(1/x)+log(1/y)), freq=FALSE, breaks=1e5, ylim=s.ylim, xlim=c(0,10), 
     main="Z(log(1/X)+log(1/Y))", xlab="",ylab="")


fw_dens_pareto <- function(w, m, n){
  
  denom <- m*n*(n-m)*w^2
  num1 <- n^2 * (1-exp(-m*w)*(1+m*w))
  num2 <- m^2 * (1-exp(-n*w)*(1+n*w))

  (num1-num2) / denom
}

# lines(seq(s.xlim[1], s.xlim[2],0.01 ),
#       fw_dens_pareto(seq(s.xlim[1], s.xlim[2],0.01 ),
#               aa_x,
#               aa_y),
#       lwd=overlaid.dens.lwd,
#       col="magenta" 
# )

lines(seq(0, 10,0.01 ),
      fw_dens_pareto(seq(0, 10,0.01 ),
              aa_x,
              aa_y),
      lwd=overlaid.dens.lwd,
      col="magenta" 
)


plot(NA,NA,axes=F,xlim=c(0,1),ylim=c(0,1),xlab="",ylab="")
```
  


  * What is the opposite of "concentration" -- bleeding out?  Multiplying the convolution by a random number bigger than 1 disperses it.




